---
title: "Immigration, Remittances, and Labor Market Outcomes"
subtitle: "An exploratory development and labor-economics project using World Bank data"
format: html
---

```{r}
#| message: false
library(WDI)
library(tidyverse)
library(ggplot2)
library(scales)
library(knitr)

fetch_indicator <- function(code, value_name, include_extra = FALSE, retries = 3) {
  for (attempt in seq_len(retries)) {
    result <- try(
      WDI(
        indicator = code,
        start = 2000,
        end = 2023,
        extra = include_extra
      ),
      silent = TRUE
    )

    if (!inherits(result, "try-error") && code %in% names(result)) {
      if (include_extra) {
        return(result |>
          rename(!!value_name := all_of(code)))
      }

      return(result |>
        select(country, iso2c, iso3c, year, all_of(code)) |>
        rename(!!value_name := all_of(code)))
    }

    Sys.sleep(2 * attempt)
  }

  stop(paste("Unable to download indicator:", code))
}

migration_data <- fetch_indicator("SM.POP.NETM", "net_migration", include_extra = TRUE)
remittance_data <- fetch_indicator("BX.TRF.PWKR.CD.DT", "remittances_received")
unemployment_data <- fetch_indicator("SL.UEM.TOTL.ZS", "unemployment_rate")
gdp_data <- fetch_indicator("NY.GDP.PCAP.CD", "gdp_per_capita")

analysis_data <- migration_data |>
  left_join(remittance_data, by = c("country", "iso2c", "iso3c", "year")) |>
  left_join(unemployment_data, by = c("country", "iso2c", "iso3c", "year")) |>
  left_join(gdp_data, by = c("country", "iso2c", "iso3c", "year")) |>
  filter(!is.na(region), !is.na(income))

latest_data <- analysis_data |>
  filter(
    year == 2023,
    !is.na(remittances_received),
    !is.na(unemployment_rate),
    !is.na(gdp_per_capita)
  )
```

## Summary

This project explores migration-related economic indicators across countries, with a focus on remittance inflows, unemployment, and development outcomes.

Using the World Bank's World Development Indicators, the analysis compares countries in the most recent year available and examines how migration-related variables vary across income groups over time. The goal is to create a clear, reproducible starting point for thinking about migration as a labor-market and development issue.

## Cross-Country Comparison

```{r}
#| fig-cap: "Remittances and Unemployment Across Countries (2023)"

ggplot(latest_data, aes(x = remittances_received, y = unemployment_rate)) +
  geom_point(aes(color = region, size = gdp_per_capita), alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.7) +
  scale_x_log10(labels = label_number(scale_cut = cut_short_scale())) +
  scale_size_continuous(name = "GDP per Capita", range = c(2, 9)) +
  labs(
    title = "Remittance Inflows and Labor-Market Outcomes",
    subtitle = "Bubble size represents GDP per capita",
    x = "Personal Remittances Received (log scale)",
    y = "Unemployment Rate (%)",
    color = "Region"
  ) +
  theme_minimal()
```

## Income Group Trends

```{r}
#| fig-cap: "Average Net Migration by Income Group (2000-2023)"

migration_trends <- analysis_data |>
  filter(!is.na(net_migration), !is.na(income)) |>
  group_by(year, income) |>
  summarise(avg_net_migration = mean(net_migration, na.rm = TRUE), .groups = "drop")

ggplot(migration_trends, aes(x = year, y = avg_net_migration, color = income)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Net Migration Trends by Income Group",
    x = "Year",
    y = "Average Net Migration",
    color = "Income Group"
  ) +
  theme_minimal()
```

## Snapshot Table

```{r}
#| tbl-cap: "Selected 2023 Country Snapshot"

latest_data |>
  select(country, region, income, remittances_received, unemployment_rate, gdp_per_capita) |>
  arrange(desc(remittances_received)) |>
  slice_head(n = 12) |>
  mutate(
    remittances_received = round(remittances_received, 0),
    unemployment_rate = round(unemployment_rate, 1),
    gdp_per_capita = round(gdp_per_capita, 0)
  ) |>
  knitr::kable(
    col.names = c(
      "Country",
      "Region",
      "Income Group",
      "Remittances (US$)",
      "Unemployment (%)",
      "GDP per Capita (US$)"
    ),
    format = "markdown"
  )
```

## Interpretation

This project is meant for careful exploratory comparison rather than strong causal claims. Differences in remittance inflows, migration balances, and labor-market outcomes can reflect institutions, conflict, demographic structure, opportunity, and broader development conditions, making the topic well suited for deeper economics analysis.
